---
title: "Reproducible figures"
output:
    html_document: default
  pdf_document: default
date: "2024-12-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#Load Packages
library(tinytex)
library(palmerpenguins)
library(ggplot2)
library(janitor)
library(dplyr)
library(svglite)

#Load functions

source("Functions/cleaning.R")
source("Functions/plotting.R")

```

*The following is a template .rmd RMarkdown file for you to use for your homework submission.*

*Please Knit your .rmd to a PDF format or HTML and submit that with no identifiers like your name.*

*To create a PDF, first install tinytex and load the package. Then press the Knit arrow and select "Knit to PDF".*

## QUESTION 01: Data Visualisation for Science Communication

*Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. **Do not make a boxplot**.*

*Use the following references to guide you:*

-   [*https://www.nature.com/articles/533452a*](https://www.nature.com/articles/533452a){.uri}
-   [*https://elifesciences.org/articles/16800*](https://elifesciences.org/articles/16800){.uri}

*Note: Focus on visual elements rather than writing misleading text on it.*

### a) Provide your figure here:

```{r bad figure code, echo=FALSE}

Bad_plot <- ggplot(penguins, aes(x = species, y = body_mass_g)) +
  geom_bar(stat = "summary", fun = "mean", fill = "blue") +
  coord_cartesian(ylim = c(3000, 6000)) + # Truncate the y-axis
  labs(title = "Average Body Mass by Penguin Species",
       y = "Body Mass (g)", x = "Species") +
  theme_minimal()

Bad_plot
# There is no need to provide the code for your bad figure, just use echo=FALSE so the code is hidden. Make sure your figure is visible after you knit it.  

```

### b) Write about how your design choices mislead the reader about the underlying data (200-300 words).

This figure misleads the reader in a number of ways. Firslty, it is using the mean mass for each species of penguin. This hides the variation that exists within the masses of each species of penguin. By not indicating the range of data for each species, the reader can not gain any information about the spread of the data and may assume it is more uniform than it is in reality. Secondly, i have truncated the y axis to a range of 3000-6000.This misleads the reader by suggesting that Gentoo penguins have a much higher mass than the other species of penguin. By using the same colour for each species it also makes it more difficult for the reader to intuitively identify what the differences shown in the plot represent. Lastly, because the figure does not show any individual datapoints, it is impossible for the reader to know the sample size for each of these penguin species.

*Include references.*

------------------------------------------------------------------------

## QUESTION 2: Data Pipeline

*Write a data analysis pipeline in your .rmd RMarkdown file. You should be aiming to write a clear explanation of the steps, the figures visible, as well as clear code.*

*Your code should include the steps practiced in the lab session:*

-   *Load the data*

-   *Appropriately clean the data*

-   *Create an Exploratory Figure (**not a boxplot**)*

-   *Save the figure*

-   ***New**: Run a statistical test*

-   ***New**: Create a Results Figure*

-   *Save the figure*

*An exploratory figure shows raw data, such as the distribution of the data. A results figure demonstrates the stats method chosen, and includes the results of the stats test.*

*Between your code, communicate clearly what you are doing and why.*

*Your text should include:*

-   *Introduction*

-   *Hypothesis*

-   *Stats Method*

-   *Results*

-   *Discussion*

-   *Conclusion*

*You will be marked on the following:*

### a) Your code for readability and functionality

### b) Your figures for communication

### c) Your text communication of your analysis

*Below is a template you can use.*

------------------------------------------------------------------------

## Introduction

The Palmer.penguins package includes data collected from three penguin species observed in the Palmer Archipelago, Antarctica, that were collected by Dr. Kristen Gorman with Palmer Station LTER. This data set includes a large number of measurements regarding the morphological characteristics of these penguins. In this analysis, i am interested in investigating the relationship between the body mass of penguins and their flipper length. This is of interest as may provide insight into the functional morphology of penguin anatomy. Flipper morphology will impact swimming efficiency, and body mass will impact bouyancy and energy expenditure. This analysis will exploring the relationship between these features and may provide insight into morphological constraints in penguins.

Before carrying out this analysis, however, it is important to prepare the data for the analysis i am planning to do.

### Loading and preparing the data for analysis

firstly, it is important to load the correct data and format it correctly.

```{r Data Exploration}

#Firstly i save the raw penguin data as a csv.This data comes from the PalmerPenguins package. This will not be overwrite to keep a record of the original data provided

write.csv(penguins_raw, "Data/data_raw.csv")

#View the Column names and the table of the raw data 

names(penguins_raw)
head(penguins_raw)

#from viewing the raw data we can see it is not well set up for analysis, for example it contains spaces.

```

### Cleaning data

I then clean the raw penguin data to make it easier to analyse. This uses functions defined in the cleaning.R file that can be found in the functions folder.

```{r}

#Clean the data by removing NAs, shortening species names and formatiing column names to be used in R. 

penguins_clean <- penguins_raw %>%
    clean_column_names() %>%
    shorten_species() %>%
    remove_empty_columns_rows()

# We can then view the new column names of the cleaned data
names(penguins_clean)

```

### Save clean data

This data frame is much better suited to analysis, we can then save this cleaned up version.

```{r}
# Save cleaned data as a csv.
write.csv(penguins_clean, "Data/penguins_clean.csv")

head(penguins_clean)
```

Having saved the new data frame we can plot an exploratory plot.

## Exploratory plot

An exploratory plot shows the raw data without any analysis or a model fitted. In this case, the most useful exploratory plot is a scatter plot that may show whether there appears to be any correlation between these two continuous variable.

### Subset data

firstly, we subset the data so that it only contains the information we are interested in. In this case the only data we are interested in is the flipper length and body mass of each penguin.

```{r Data Exploration}
#filter clean data to only show mass and flipper length

Mass_flipper_data <- penguins_clean %>%
  subset_columns(c("body_mass_g", "flipper_length_mm")) %>%
  remove_NA()

#view this new data frame
head(Mass_flipper_data)

#create a new csv with this subset of the data
write.csv(Mass_flipper_data, "Data/Mass_flipper_data.csv")


```

### Generate an exploratory plot

```{r}

#Plot the exploratory table, this uses a function from the plotting.R file found in the functions folder

 Mass_flipper_plot <- Plot_explor(Mass_flipper_data)
 Mass_flipper_plot
 
```

From this exploratory figure we can see that there appears to be a strong positive correlation between body mass and flipper length.

### Save figure

```{r}
#Save this figure as an SVG file

svglite("Figures/Mass_flipper_plot.svg",
        width = 5.9, height = 5.9)
Mass_flipper_plot
dev.off()

```

## Hypothesis

My hypothesis here is that flipper length is correlated with body mass. The null hypothesis in this analysis is that there is no relationship between flipper length and body mass and the alternative hypothesis is that there is a correlation between the two measures.

### Statistical Methods

To identify if there is a correlation between flipper length and body mass, i can create a linear regression model. This can be done by using the lm() function which will fit a linear regression model to the data and give me the coefficients and the p values associated with these coefficients.

```{r Statistics}

#I create a linear model using the lm() function:

linear_model <- lm(flipper_length_mm ~ body_mass_g, Mass_flipper_data)

# Extract coefficients and R-squared
coefficients <- summary(linear_model)$coefficients
r_squared <- summary(linear_model)$r.squared

# Combine coefficients with R-squared
output <- list(
  Coefficients = coefficients,
  R_squared = r_squared
)

# Print the output
output

```

From the summary table we can see that there is positive correlation between flipper length and body mass. The Estimate for the Y intercept is 1.367e+02 and the estimate for the slope is 1.528e-02 We can see form the r-squared value that body mass can predict 75.9% of variation in flipper length, making it biologically significant. The p values for both of these estimates is highly significant at \<2e-16

### Validation of assumptions.

In order to use this linear model, we need to ensure some assumptions are met. the assumptions for this model is that the residuals of this data are normally distributed and that the variance is roughly equal for all values of x. We can visually assess whether these assumption are met by plotting out the graphs below.

```{r}
#Shapiro Wilks test for normality
plot(linear_model, which = 2)

#Breusch-Pagan test to test for Homoscedasticity:
plot(linear_model, which = 1)
```

From visual inspection, we can confirm the assumptions stated above.

## Results & Discussion

### Results figure

We can use the results from this linear model to add our model to the graph showing the penguin data. We can also add 95% confidence intervals which can be seen in the pink shaded area around the linear regression line.

```{r Plotting Results}

#plot the raw data with the linear regression line and the 95% confidence intervals. The function for this can be found in plotting.R
Results_figure <- Plot_point(Mass_flipper_data)
Results_figure


```

### Save results figure

```{r}
#save results figure as an svg file. 

svglite("figures/Results_figure.svg",
        width = 5.9, height = 5.9)
Results_figure
dev.off()

```

### Conclusion

My analysis shows that there is a positive correlation between flipper length and body mass in the penguins from the palmer's penguins data set. Body mass predicts 75.9% of variation in flipper length. Further study could investigate whether this correlation exists or is different between the three species present in the data set.

------------------------------------------------------------------------

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*
